Compiling Python Modules to Native Parallel Modules Using Pythran and OpenMP Annotations

### 2013-09-08 doc/papers/sc2013/paper.tex
Minor grammar fix to SC2013

    High Performance Computing users traditionally rely on low-level,
    compiled language such as C or
    FORTRAN to perform computation-intensive tasks. As a consequence, it is a
    common situation to have High Performance Computing application written in a high-level language
    such as Python, calling native routines for computate-intensive tasks. To
    improve development speed and reduce maintenance costs, using a higher-level
    language like Python seems attractive. While it is usually associated with
    low performance, several solutions such as Cython, Numba, Parakeet or
    Pythran offer to automatically or semi-automatically turn Python functions
    into native ones.

@@

Parallel computations are supported by the Python standard library through the
\texttt{multiprocessing} module. It spawns several interpreters that
communicate through Inter Process Communication (IPC), using Python built-in
object serialization. This approach is only viable for computation intensive
independent tasks, since the communication and synchronization overheads are
much greater than a light-weight threaded approach.

@@

An opposite approach consists in using the host language ---herein Python--- to
describe both parts of the system, i.e.\ the hybrid and the native. An automated
tool performs the translation to native code of a specific part of the
application, generally the computation-intensive one where parallelism has been
expressed in some ways.
Therefore, developers not familiar with lower level languages or not eager to
invest the additional development time can still benefit from a fair performance
boost.
This approach is the subject of many studies that can be classified according to
their compatibility with the host language.

@@

%% For the two paragraph below. it is not specified if openmp annotations were
% added by hand. It is misleading since the previous paragraph said
% "The code itself is not modified, only the compilation flag."
Figure~\ref{fig:pb-julia} shows the execution time of the computation of the
julia fractal. It's a highly parallel, computate-intensive code. Pythran and
Cython's version have almost the same execution time, but activating OpenMP
yields an extra $\times 3.95$ speedup when compared to Pythran alone.

>>>>

    High Performance Computing users traditionally rely on low-level,
    compiled language such as C or
    FORTRAN to perform compute-intensive tasks. As a consequence, it is a
    common situation to have High Performance Computing application written in a high-level language
    such as Python, calling native routines for compute-intensive tasks. To
    improve development speed and reduce maintenance costs, using a higher-level
    language like Python seems attractive. While it is usually associated with
    low performance, several solutions such as Cython, Numba, Parakeet or
    Pythran offer to automatically or semi-automatically turn Python functions
    into native ones.

@@

Parallel computations are supported by the Python standard library through the
\texttt{multiprocessing} module. It spawns several interpreters that
communicate through Inter Process Communication (IPC), using Python built-in
object serialization. This approach is only viable for compute-intensive
independent tasks, since the communication and synchronization overheads are
much greater than a light-weight threaded approach.

@@

An opposite approach consists in using the host language ---herein Python--- to
describe both parts of the system, i.e.\ the hybrid and the native. An automated
tool performs the translation to native code of a specific part of the
application, generally the compute-intensive one where parallelism has been
expressed in some ways.
Therefore, developers not familiar with lower level languages or not eager to
invest the additional development time can still benefit from a fair performance
boost.
This approach is the subject of many studies that can be classified according to
their compatibility with the host language.

@@

%% For the two paragraph below. it is not specified if openmp annotations were
% added by hand. It is misleading since the previous paragraph said
% "The code itself is not modified, only the compilation flag."
Figure~\ref{fig:pb-julia} shows the execution time of the computation of the
julia fractal. It's a highly parallel, compute-intensive code. Pythran and
Cython's version have almost the same execution time, but activating OpenMP
yields an extra $\times 3.95$ speedup when compared to Pythran alone.
